### 2015-10-17 submission.tex
integrated lan's commets for section1,2,3,4

\documentclass[conference]{IEEEtran}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
\else

\fi

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}
\usepackage{algorithm}
%\usepackage{algorithm2e}
\usepackage{algpseudocode}
\usepackage{booktabs}
\usepackage{flushend}
\usepackage{url}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{textcomp}
%\usepackage{subfig}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{soul}
\usepackage{verbatim}
%\usepackage[labelfont=bf]{caption}
\usepackage[font=bf,labelsep=space]{caption}

\newcommand{\TODO}[1]{\hl{TODO: #1}}
\newcommand{\NOTE}[1]{\hl{NOTE: #1}}
\newcommand{\changed}[1]{\textcolor{red}{#1}}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Study of Intra- and Inter-Job Interference with Different Job Allocations}
% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{

\IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B. Ross\IEEEauthorrefmark{2}, Zhou Zhou\IEEEauthorrefmark{1}, Zhiling Lan\IEEEauthorrefmark{1}}

\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
Illinois Institute of Technology,
Chicago, Illinois, USA 60616\\
\{xyang56, zzhou\}@hawk.iit.edu, lan@iit.edu}

\IEEEauthorblockA{\IEEEauthorrefmark{2}Mathematics and Computer Science Division, Argonne National Laboratory,
Argonne, IL, USA 60439\\
\{jenkins,rross\}@mcs.anl.gov, mmubarak@anl.gov}
}

% make the title area
\maketitle

\begin{abstract}

Network contention between concurrently running jobs on HPC systems is a primary cause of performance variability. Optimizing job allocation and avoiding network sharing are hence crucial to alleviate the potential degradation in performance. In order to do these effectively, an understanding of the interaction between applications, their communication patterns, and the system architecture are required. In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference. By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference induced by different job allocation scenarios. Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.

\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction} 
\label{sec: intro}

The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research. Supercomputers have tens of thousands of nodes  and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity. Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently.  These parallel jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources. As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.

\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/goodallocation}
        \caption{Contiguous}
        \label{fig:overview_sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/badallocation}
        \caption{Non-contiguous}
        \label{fig:overview_sub2}
    \end{subfigure}%
   \caption{Multiple jobs running concurrently with different allocations. Each job is represented by a specific color. a) shows the effect of contiguous allocation, which reduce the inter-job interference. b) shows non-contiguous allocation, which may introduce both intra and inter-job interferences. }
   \label{fig:overview}
\end{figure}

When submitting jobs to HPC systems, users request resources in terms of node counts and expected runtime. The batch scheduler is then responsible for dispatching them to the system. Typically, multiple jobs run concurrently, resulting in the shared use of resources, particularly network and storage. A prominent problem with this approach is network contention among concurrently running jobs. If the network resources of concurrently running jobs aren't isolated, the resulting sharing can cause communication variability that results in performance degradation for some or all such jobs \cite{abhinav-sc13}. This delay can propagate into the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}. This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resources.

On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used. The \emph{contiguous allocation} strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}. The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \cite{bgloverview}. The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies. However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible. On the other side of the coin, the \emph{non-contiguous allocation} strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality. Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation. While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes. The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \cite{abhinav-sc13}.

We envision that future HPC schedulers will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation strategies. Such a flexible mechanism would take shared resource needs of jobs into account when making allocation decisions (e.g., network). With knowledge and analysis of job communication patterns, it can be identified which jobs require network isolation and locality, and to what degree. Then, rather than allocating each job in a ``know-nothing'' manner, one may specialize allocation so that, for example, only the jobs with stringent network needs are given compact, isolated allocations, resulting in maximized utilization and minimized perceivable resource contention effects.

In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems. Torus-based networks are used on six of the top 10 supercomputers on the June 2015 Top500 lists~\cite{top500}. The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}. The K computer from Japan uses the ``Tofu'' system, which is a 6D mesh/torus topology \cite{tofu}. Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}. Although our analyses are based on torus networks, the ideas conveyed in this work is applicable to networks with different topologies.

We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns. We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment. We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations. We believe the insights presented in this work can be very useful for the design of future HPC batch job schedulers and resource managers.

%\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
\begin{comment}
The observations of this work are the following:

\begin{itemize}
    \item We study the communication behaviors and identify the communication patterns of three representative HPC applications.
    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocation strategies. And we find the relation between application's communication pattern and the cause of interference.
    \item We show a novel and flexible way to make allocation with awareness of job's communication pattern, which  is every effective to alleviate the interference between concurrently running jobs. \TODO{do we actually do this, or provide the underpinnings for such a method?}
\end{itemize}

We believe the finds and insights presented in this work would be very useful for the design of future HPC batch job scheduler and resource management module.

\end{comment}

The rest of this paper is organized as follows. Section~\ref{sec:application study} describes the three representative applications from DOE Design Forward Project for our study. Section~\ref{sec:codes} talks about the use of CODES as research vehicle for our work. Section~\ref{sec:config study} shows the performance analysis of three applications on different torus networks. Section~\ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocation strategies. Section~\ref{sec:discussion} introduces a path toward communication-pattern aware allocation strategies, given the results of our analysis. Section~\ref{sec:related_work} discusses related work. Finally, the conclusion is presented in Section~\ref{sec:conclusion}.

\section{Application Study}
\label{sec:application study}

%A parallel application usually conforms to a combination of several basic communication patterns\cite{roth}. At its different execution phases, the application's communication behavior may follow different basic patterns respectively. \textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution. Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc. In this work, we focus on the locality of application's rank-to-rank communication topology graph.

A parallel application usually conforms to a combination of several basic communication patterns \cite{roth}. At its different execution phases, the application's communication behavior may follow different basic patterns respectively. There are many profiling tools available to capture information regarding communication patterns of parallel applications~\cite{tau,mpip,scala,sst,oxbow}. They can provide information such as the percentage of different MPI operations of the applications, communication topology, the amount of data transferred between processes.

There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns. In this work, we select three representative applications from the DOE Design Forward Project. Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications. We believe that the communication patterns of these applications are representative of a wide array of applications running on leadership-class machines. Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps. The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.

% MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}. Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures. We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter. Their representative communication patterns that are commonly seen in the HPC system workload. The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{designforwardwebpage}.

\subsection{AMG}
\label{sec:amg}

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.25\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
        \caption{}
        \label{fig:amg-communication-topology}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.24\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
        \caption{}
        \label{fig:amg-data-trans}
    \end{subfigure}
    \caption{AMG. (a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in AMG. (b) The amount of data sent/received by each rank in AMG. Y-axis labels the amount of data transferred in MB, x-axis is the rank. }
\end{figure}

The Algebraic MultiGrid Solver, or AMG, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages. It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}. The dominant communication pattern is regional communication with decreasing message size for different parts of the multi-grid v-cycle.

Figure~\ref{fig:amg-communication-topology} shows the communication topology of a small scale AMG execution with 216 MPI Ranks. Note that the dominant communication pattern of the application doesn't change with scale. We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks. Applications with similar pattern include PARTISN \cite{partisn} and SNAP \cite{snap}.

Figure~\ref{fig:amg-data-trans} shows the amount of data each rank transferred in the example execution. The data transferring of each rank is between 7MB to 16MB, corresponding to whether the rank is on a boundary or not.

\subsection{Crystal Router}
\label{sec:crystalrouter}

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.25\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
        \caption{Communication Topology}
        \label{fig:cr-communication-topology}
    \end{subfigure}
    ~
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
        \caption{Data Volume}
        \label{fig:cr-data-trans}
    \end{subfigure}
    \caption{CrystalRouter. (a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in CrystalRouter. (b) The amount of data sent/received by each rank in CrystalRouter. Y-axis labels the amount of data transferred in MB, x-axis is the rank. }
\end{figure}

The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \cite{nek5000}. It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos. CrystalRouter demonstrates the ``many-to-many'' communication pattern through a scalable multi-stage communication process. The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.

The collective communication in CrystalRouter utilizes a recursive doubling approach. Ranks in CrystalRouter conform to a $n$-dimensional hypercube and recursively split into ($n$-1)-dimensional hypercubes, with communication occurring along the splitting plane. The pattern of this communication can be found in Figure~\ref{fig:cr-communication-topology}. By nature of the logarithmic splitting process, a substantial portion of the communication occurs in small neighborhoods of ranks. CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication, and shares similarities with most MPI collective communication implementations.

Figure~\ref{fig:cr-data-trans} shows the amount of data each rank transferred in CrystalRouter, which ranges between 20MB to 36MB. This is due to the rank counts not matching one-for-one with points in the hypercube.

\subsection{MultiGrid}
\label{sec:multigrid}

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.25\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
        \caption{Communication Topology}
        \label{fig:mg-communication-topology}
    \end{subfigure}
    ~
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
        \caption{Data Volume}
        \label{fig:mg-data-trans}
    \end{subfigure}
    \caption{MultiGrid. (a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in MultiGrid. (b) The amount of data sent/received by each rank in MultiGrid. Y-axis labels the amount of data transferred in MB, x-axis is the rank. }
\end{figure}

MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \cite{boxlib}. MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle. It is widely used for structured grid physics packages.

Figure~\ref{fig:mg-communication-topology} shows the communication topology of MultiGrid with 125 ranks. We can see intensive communication along the diagonal that resembles nearest neighbor communication, similar to AMG. However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, challenging the maximization of communication locality with respect to rank. In this sense, it can be considered a ``many-to-many'' pattern. Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in \cite{boxlib}.

Figure~\ref{fig:mg-data-trans} shows the amount of data transferred between ranks in MultiGrid. For this miniapp, The data transferred per rank is relatively small and stable, between 2.8MB to 2.9MB.

\section{Research Vehicle}
\label{sec:codes}

It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context. One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by users. Even the system administrator is not authorized to make that change. Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference and then compare the results with those in the presence of interference. Therefore, we resort to simulation for this work.

A simulation toolkit named CODES enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}. CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}. CODES support both torus and dragonfly network with high fidelity flit-level simulation. CODES has this network workload component that is capable of conducting trace-driven simulations. It can take real MPI application traces generated by SST DUMPI \cite{sst} to drive CODES network models. In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems. Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.

\section{Configuration Study}
\label{sec:config study}

In this section, we study the communication behavior of three applications on torus networks with different dimensionality and bandwidth configurations. Torus networks are $k$-ary $n$-cubes, with $k^{n}$ nodes in total arranged in an $n$-dimensional grid having $k$ nodes in each dimension. Each node has $2\times n$ direct linked neighbor nodes with wraparound. In these and all following experiments, non-adaptive routing is used. While making specific recommendations as to an ``optimal'' network configuration for a given application is out of the scope of this study, the experiments do highlight the nuances inherent in design choices such as network diameter and bandwidth planning.

We perform three sets of experiments. Each looks at per-rank distribution of performance of AMG, CrystalRouter and MultiGrid on a set of 2048-node torus models, one with $n=3$ ($16^{2} \times 8$), one with $n=5$ ($8\times 4^{4}$), and one with $n=7$ ($4^{4} \times 2^{3}$). Performance in this context is defined by the total message transfer time for each sending rank, which we are able to precisely measure due to the simulation environment providing us with an absolute notion of time. Note that we simply perform linear rank assignment, one per node -- later sections investigate rank and node mapping effects. In the first, shown in Figure~\ref{fig:dimensionality-study}, we keep the link bandwidth constant at 2GiB/s in each direction, corresponding to Blue Gene/Q systems~\cite{bgq}, leading to aggregate node bandwidths of 12GiB/s, 20GiB/s, and 28GiB/s for the 3D, 5D, and 7D tori, respectively. In the second, shown in Figure~\ref{fig:bandwidth-study}, we fix the torus dimensionality to 5D. In the third, shown in Figure~\ref{fig: bandwidth-time-box}, we instead fix the per-node aggregate bandwidth (and thus the system aggregate bandwidth) and modify the torus dimensionality, resulting in per-link bandwidths shown in Table~\ref{tab: fix-bandwidth}.

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/dimenstudy/amg_box}
        \caption{AMG}
        \label{fig:dimen-amg}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/dimenstudy/cr_box}
        \caption{CrystalRouter}
        \label{fig:dimen-cr}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/dimenstudy/mg_box}
        \caption{MultiGrid}
        \label{fig:dimen-mg}
    \end{subfigure}%
   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network, with constant per-link bandwidth.}
   \label{fig:dimensionality-study}
\end{figure*}

In the fixed link bandwidth experiment, shown in Figure~\ref{fig:dimensionality-study}, the data transfer time of the three applications are significantly reduced as the dimensionality increases, so the applications can easily make use of the increased dimensionality (lower hops on average) and/or aggregate bandwidth (more links being concurrently used). Interestingly, AMG has the greatest degree of outliers compared to the others. We expect this to be a combination of non-optimal rank mappings causing a greater degree of variance due to the highly regular communication pattern and the distribution of message sizes. MultiGrid seems to benefit the most in terms of reducing the spread of the time distribution, though the median message time remains roughly within a factor of two. The time distribution in CrystalRouter seems to accrue the most benefits, likely due to the combination of cross-cutting and localized per-rank traffic.

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/bandwidthstudy/amg_bw_box}
        \caption{AMG}
        \label{fig:bdwstudy-amg}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/bandwidthstudy/cr_bw_box}
        \caption{CrystalRouter}
        \label{fig:bdwstudy-cr}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/bandwidthstudy/mg_bw_box}
        \caption{MultiGrid}
        \label{fig:bdwstudy-mg}
    \end{subfigure}%
   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 5D torus network with 2GiB/s, 4GiB/s and GiB/s direct link bandwidth. }
   \label{fig:bandwidth-study}
\end{figure*}

In the fixed dimensionality experiment, shown in Figure~\ref{fig:bandwidth-study}, the increased per-link bandwidth, expectedly, benefits each application roughly equally, reducing the time distributions linearly.

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/samebdw/amg}
        \caption{AMG}
        \label{fig:samebd-amg}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/samebdw/cr}
        \caption{CrystalRouter}
        \label{fig:samebd-cr}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/samebdw/mg}
        \caption{MultiGrid}
        \label{fig:samebd-mg}
    \end{subfigure}%
   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network with same aggregated bandwidth. }
   \label{fig: bandwidth-time-box}
\end{figure*}

\begin{table}[ht]
\begin{center}
\caption{Torus networks with same aggregated bandwidth} 
\label{tab: fix-bandwidth}
\begin{tabular}{l c c c} 
\toprule % Top horizontal line
\toprule
&\multicolumn{3}{c}{Bandwidth } \\
\cmidrule(l){2-4}
Dimension  & per-link & aggregate &\\ % Column names row
\midrule % In-table horizontal line
3D      & 4.67GiB/s  & 28GiB/s  & \\  % Content row 1
\midrule % In-table horizontal line
5D    & 2.8GiB/s  & 28GiB/s  &\\ % Content row 1
\midrule % In-table horizontal line
7D    & 2GiB/s & 28GiB/s & \\
\midrule
\bottomrule % Bottom horizontal line
\end{tabular}
\end{center}
\end{table}

In the fixed aggregate bandwidth experiment, with configuration in Table~\ref{tab: fix-bandwidth} and results shown in Figure~\ref{fig: bandwidth-time-box}, the results are more nuanced. At least for these applications, the benefits from increased dimensionality do not outweigh the reduction in per-link bandwidth. MultiGrid does see benefits in the increased dimensionality, particularly with respect to outliers, but has a similar median message time. CrystalRouter reduces performance as the dimensionality increases, but only by a small degree. AMG remains roughly the same, minus a dip in the maximum message time that could be due to rank-to-node mapping changes.

\section{Interference Analysis}
\label{sec:interference}

Communication variability due to network sharing can cause application performance degradation. \textbf{\emph{Intra-job interference}} refers to the network contention between ranks within each application. \textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.

In this section we will study both kinds of interference in the context of torus network. 
Since application's communication pattern won't change with its scale, we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks. And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.

\subsection{Intra-Job Interference Analysis}
\label{sec: introjob}

We design two sets of experiments to study the intra-job interference of each application. In the first, shown in Figure \ref{fig:shapestudy}, we assign each application with allocation in three different shapes.  In the second, shown in Figure \ref{fig:diffmap}, we assign each application with 3D-balanced cube and use different mapping strategies for rank-to-node mapping.

In the allocation shapes experiment, we select three possible shapes can can commonly seen on a 3D torus network. They are 3D balanced-cube, 3D unbalanced-cube, 2D mesh.

3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation. There are some research work like \cite{leung} \cite{abhinav-sc13} try to prove that compact allocation can guarantee job with better performance. In this experiment, 3D balanced-cube is the most compact allocation on a 3D torus network.

3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on system with asymmetric networks. The Blue Waters, a Cray XE6/XK7 system network is 3D torus with Gemini routers. The network connections in the \emph{y} direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions. In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation is always start from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.

2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus. 2D mesh is a very common allocation shape in torus network for both contiguous and non-contiguous allocation strategies. For example, Cray Application Level Placement Scheduler indexes the nodes in torus network into a list and make allocation by simply works off the this list \cite{carl-cug}. When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh. The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/allocshape/allocation}
        \caption{ }
        \label{fig:cont_sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit16}
        \caption{ }
        \label{fig:noncont_sub1}
    \end{subfigure}%
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit8}
        \caption{ }
        \label{fig:noncont_sub2}
    \end{subfigure}%
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit2}
        \caption{ }
        \label{fig:noncont_sub3}
    \end{subfigure}%
    \caption{(a) Contiguous allocation in three different shapes. ``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube. (b)-(d) Non-contiguous allocation. Each job is represented by a specific color. The nodes assigned to different jobs are interleaved, the size of allocation unit are 16, 8 and 2.}
\end{figure*}

Figure \ref{fig:shapstudy-amg} shows that for AMG, who conforms to 3D ``Nearest Neighbor" communication pattern, spent less time when being assigned with 3D-unbalanced allocation than 3D-balanced. And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}. Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message between ranks in MultiGrid. CrystalRouter has with both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.

There are lots of research work try to design fancy allocation algorithms to provide application with most compact allocation. However, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application. Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern. Application with dominant communication pattern like ``Nearest Neighbor" won't fully utilize the compact allocation and could get even better performance on a relaxed allocation shape.

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/amg_box}
        \caption{AMG}
        \label{fig:shapstudy-amg}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/cr_box}
        \caption{CrystalRouter}
        \label{fig:shapstudy-cr}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/mg_box}
        \caption{MultiGrid}
        \label{fig:shapstudy-mg}
    \end{subfigure}%
   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 3D-unbalanced and 2D allocation.}
   \label{fig:shapestudy}
\end{figure*}

\TODO{This par is OUTDATED}. In non-contiguous allocation systems, the nodes assigned to each job could be scatter anywhere in the system. Some HPC systems predefine the minimum number of nodes consist of the allocation unit. The nodes within each unit are contiguous, and units assigned to each job may have arbitrary locations. We assign each application with a non-contiguous set of allocation units, refers to a small number(like 4) of adjacent nodes as shown in Figure \ref{fig:noncont_sub1}. We compare application's performance when it runs with contiguous allocation and non-contiguous allocation with unit size of 16, as shown in Figure \ref{fig: cont-time-box}.

\TODO{This par is OUTDATED} Figure \ref{fig:cont-amg} shows that AMG is insensitive to the allocation's contiguity. Non-contiguous allocation can serve AMG equally well as contiguous allocation, as long as the allocation unit in non-contiguous allocation is big enough to preserve the locality of AMG. CrystalRouter shows preference in choosing between contiguous and non-contiguous allocation. As shown in Figure \ref{fig:cont-cr}, there is obvious degradation in terms data transfer time when CrystalRouter running with non-contiguous allocation. The global data transfer in CrystalRouter will take more hops on non-contiguous allocation, thus result in longer communication time. When it comes to MultiGrid shown in Figure \ref{fig:cont-mg}, this degradation is even worse. Since the global data transfer take a great portion of MultiGrid ``Many-to-Many" communication pattern.

\TODO{This par is OUTDATED} The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different allocations. Non-contiguous allocation won't aggravate intra-job interference as long as the unit size can preserve the locality in application's communication topology graph.

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/diffmapping/amg}
        \caption{AMG}
        \label{fig:diffmap-amg}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/diffmapping/cr}
        \caption{CrystalRouter}
        \label{fig:diffmap-cr}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/diffmapping/mg}
        \caption{MultiGrid}
        \label{fig:diffmap-mg}
    \end{subfigure}%
   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D-balanced allocation with different mapping strategies.}
   \label{fig:diffmap}
\end{figure*}

\subsection{Inter-Job Interference Analysis}
\label{sec: interjob interference study}

%We simulate three application running concurrently with different allocation to analysis the interference between them. First, we assign each application with contiguous allocation side by side, and compare their performance with each application running exclusively. For AMG, sharing the network with other application won't affect its performance, since all the data transfer are between neighboring ranks. CrystalRouter and MultiGrid will suffer prolonged communication time when sharing network with others. The results are shown in Figure \ref{fig: concurrent-cont-time-box}.
Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}. Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than system with contiguous allocation. Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.

We assign each application with non-contiguous allocation and run them concurrently on the same network. The allocation unit belongs to different jobs are interleaved as shown in Figure \ref{fig:noncont_sub1}. The unit size is critical to the application's performance. Figure \ref{fig: 3apps interjob study} shows the results of each application data transfer time with different allocation unit size.

The data transfer time of AMG in Figure \ref{fig:interjob-amg-box} keeps stable between allocation unit size of 16 (big) and 8 (medium), since the locality of AMG is 6-rank based. When the unit size reduce to 2 (small), AMG suffers prolong data transfer time by about 10\%. CrystalRouter is more sensitive to allocation unit size. The best unit size would be the one big enough to accommodate its local neighborhood. Figure \ref{fig:interjob-cr-box} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16. When the unit size reduce to 2-node, the communication become less efficiency and takes about 15\% more times for transferring data.

The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg-box}. This is because even big allocation unit size like 16-node will still fail to preserve MultiGrid's ``Many-to-Many" pattern. The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg-box} . As the unit size decreases, the data transfer time keep growing, but in a slow way, which means in terms of preserving the locality of MultiGrid, allocation unit of size 16, 8 and 2 serve equally bad.

The unit size in non-contiguous allocation policy should be chosen according to application's communication pattern. Inter-job interference is inevitable in non-contiguous based systems, but unit size that big enough to preserve the neighborhood communication of the application will alleviate such interference and improve job performance.

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/inter-job/amg}
        \caption{AMG}
        \label{fig:interjob-amg-box}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/inter-job/cr}
        \caption{CrystalRouter}
        \label{fig:interjob-cr-box}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/inter-job/mg}
        \caption{MultiGrid}
        \label{fig:interjob-mg-box}
    \end{subfigure}%
   \caption{Inter-job interference study. ``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation. To study the impact of non-contiguous allocation on inter-job interference, applications running concurrently with interleaved allocation of different unit sizes, which are 16-node, 8-node, 2-node. }
   \label{fig: 3apps interjob study}
\end{figure*}

\subsection{Summary}
\label{sec:summary}

Based on our comprehensive study, we can get following observations:
\begin{itemize}
    \item The compact allocation may not guarantee the best performance for every application. 
    \item The performance of application conforms to Nearest Neighbor communication pattern keeps relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefers compact allocation. 
    \item The allocation unit size should be decided according to application's communication pattern. Unit size that big enough to preserve the neighborhood communication of the application will result in better performance. 
\end{itemize}

\section{Discussion}
\label{sec:discussion}

%HPC system usually accommodate many parallel jobs running concurrently. Each job gets the required number of computing nodes by system resource manager. The allocated nodes for each job communicate with each other through system interconnect network in a sharing way, which cause interference as a negative side-effect. This interference can introduce performance degradation to both job and system, such as prolonged job running time and reduced system throughput.

The observations we obtained by analyzing the intra- and inter-job interference can provide insights for the design of a smart and flexible job allocation strategy. By scrutinizing job's communication behavior, we can identify job's dominant communication pattern and pinpoint the ``neighborhood communication". With such knowledge about jobs' communication patterns, we can analyze the possible interference between jobs and take precautions to alleviate the negative effect when making allocation decisions.

When an application with dominant communication that is intensive ``Many-to-Many" being submitted to the system, it should be granted with compact node allocation and exclusive network provision. The compact allocation can guarantee the shortest pair-wise distance between all the ranks, and make the data transfer between ranks take less hops. On the other hand, the exclusive network provision will prevent other adjacent applications from sharing network resources, thus eliminate the performance degradation due to interference.

Not every application's preferable resource are compact node allocation and exclusive network provision. As we found in our comprehensive analysis, applications whose dominant communication patterns contain intensive ``neighborhood communication" like ``Nearest Neighbor" won't benefit from compact node allocation and exclusive network provision. Such applications can run with non-contiguous node allocation without performance degradation, as long as the allocation unit can accommodate the ``neighborhood communication".

The allocation unit size shouldn't be fixed. Instead, the new allocation strategy should choose the proper unit size based on the scale of ``neighborhood communication" of each job. The best unit size should be neither too big nor too small, jut perfectly fit for the size of that ``neighborhood". Big unit size won't be fully utilized and results in fragmentation. Small unit size won't be able to accommodate the ``neighborhood communication" and makes the intra-job communication less efficient.

The advantage of making allocation with consideration about job's communication pattern is obvious. First, compared with contiguous allocation policy, the new allocation strategy is more flexible, it doesn't require the system to provide a big contiguous partition that can accommodate the whole application, just small set of compact nodes enough for all the "local community" in the application. On the other hand, the new allocation strategy won't destroy the locality of application's communication pattern. On the contrary, the small compact node set (allocation unit) provided for the application's  ``local communities"  will preserve the communication locality in the maximum extent.

%The off-line study of job's communication traces could be helpful to develop better allocation policies for batch scheduler to make on-line scheduling/allocating decisions. This is due to the set of applications running on those leadership computing facilities is usually stable, and the workloads from those HPC systems is usually has great repetitiveness. We analyzed the workload on Mira, an IBM Blue/Gene Q system at Argonne Leadership Computing Facility, found that a large portion of its workload are repetitive. Figure \ref{fig: repetitiveness of Mira} shows the 5 of most active users with repetitive submissions. The collected jobs information from history can be very useful for workload prediction and system optimization.
%
%
%\begin{figure}[h!] 
%  \centering
%  \includegraphics[width=0.38\textwidth]{figs/repe-mira}
%   \caption{The repetitiveness of Mira's workload. }
%   \label{fig: repetitiveness of Mira}
%\end{figure}

\section{Related Work}
\label{sec:related_work}
%\textbf{P1: tools for monitoring and profiling}\\
There are some tools available for system monitoring and application profiling on HPC systems. Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces. However, recognizing the patterns about application's communication from those traces requires lots of extra effort. Some researchers work on the recognition and characterization of parallel application communication patterns. Oak Ridge National Laboratory has this ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of 12 scientific applications and benchmarks \cite{oxbow}. In their recent work \cite{roth}, they demonstrate a new approach to automatically characterizing a parallel application's communication behavior.

%\textbf{P2: communication characteristics}\\
Some research institutes put great effort in characterization of scientific applications. The Department of Energy initiate this design forward project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{designforwardwebpage}. In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload. A joint project named CORAL from Oak Ridge, Argonne and Livermore provide a series of benchmarks to represent DOE workloads and technical requirements \cite{coral}. The CORAL project includes scalable science benchmarks, throughput benchmarks, data centric benchmarks, skeleton benchmarks and Micro benchmarks.

%\textbf{P3: Interference between concurrently running jobs}\\
The interference between concurrently running jobs on HPC system have been identified as major culprit for job's performance variability. Abhinav et al. found that concurrently running applications on HPC can cause interference to each other, and cause communication time varied from 36\% faster to 69\% slower. Such interference could come from Operating System noise, shape of the allocated partition and mainly other running jobs that sharing network resources \cite{abhinav-sc13}. Skinner et al. found that  there is a 2-3$\texttimes$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs \cite{skinner}. Rosenthal et al. found there is limited benefit for many applications by increasing network bandwidth. The applications that send mostly small messages or larger messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth \cite{rosenthal}.

%\textbf{P4: Job allocation(little bit)}\\
There are some research work focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs. Hoefler et al. propose to use performance modeling techniques to analysis factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}. However, as the scale of HPC continue grows, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools. Bogdan et al provide a set of guidelines how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}. Dong et al describe IBM Blue Gene/Q's 5D torus interconnect network \cite{Dong-SC11}. They developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network.

Our work is different from all these works in the following ways. First, we focus on the dominant communication patterns rather than any specific application. We believe this can provide a guideline for other research work is the area. Secondly, we explored the inter-job interference between concurrently running jobs, while similar work such as \cite{abhinav-sc13} only focus on the single application's performance degradation due to network contention. Finally, we explored the impact of different allocation strategies to job's communication behavior. We identified the optimal allocation strategy for each application with specific dominant communication pattern. Based on our comprehensive exploration, we claim that better allocation strategy should take job's communication pattern into consideration for allocation decision making.

\section{Conclusions}
\label{sec:conclusion}
In this work, we study the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid. Each application has distinctive communication pattern, that can be representative for a group of jobs in HPC workload. We use a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network. The torus network provided by CODES has good fidelity and scalability. We analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations. We found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.

We also analysis the intra- and inter-job interference by simulating three applications running on 3D torus network. Based on our comprehensive experiments, we got three observations. 1) The compact allocation can not guarantee the best performance for every application. 2) The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefer compact allocation. 3) The allocation unit size should be decided according to application's communication pattern. Unit size that big enough to preserve the locality of the application will result better performance.

We believe that our finding in this work can provide guidance for HPC resource management to make flexible job allocations. Rather than use pre-defined partitions and reckless non-contiguous allocation, future HPC system should assign each job with preferable resources based on job's communication pattern.

\section*{Acknowledgment}
\label{sec: ack}
The work at Illinois Institute of Technology is supported in part by U.S. National Science Foundation grants CNS-1320125 and CCF-1422009. This work is also supported by the U.S. Department of Energy, Office of Science, Advanced Scientific Computing Research, under Contract DE-AC02-06CH11357.

\bibliographystyle{IEEEtran}
\bibliography{./reference}

\vspace{5\baselineskip}
The submitted manuscript has been created by UChicago Argonne, LLC, Operator of Argonne National Laboratory ("Argonne").  Argonne, a U.S. Department of Energy Office of Science laboratory, is operated under Contract No. DE-AC02-06CH11357.  The U.S. Government retains for itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide license in said article to reproduce, prepare derivative works, distribute copies to the public, and perform publicly and display publicly, by or on behalf of the Government.

\end{document}