### 2016-07-08 submission.tex
change some reference format

\documentclass[conference]{IEEEtran}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
\else

\fi

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}
\usepackage{algorithm}
%\usepackage{algorithm2e}
\usepackage{algpseudocode}
\usepackage{booktabs}
\usepackage{flushend}
\usepackage{url}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{framed}
\usepackage{textcomp}
%\usepackage{subfig}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{soul}
\usepackage{verbatim}
%\usepackage[labelfont=bf]{caption}
\usepackage[font=bf,labelsep=space]{caption}

\newcommand{\TODO}[1]{\hl{TODO: #1}}
\newcommand{\NOTE}[1]{\hl{NOTE: #1}}
\newcommand{\changed}[1]{\textcolor{red}{#1}}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Study of Intra- and Inter-Job Interference with Different Job Allocations}
% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{

\IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B. Ross\IEEEauthorrefmark{2}, Zhou Zhou\IEEEauthorrefmark{1}, Zhiling Lan\IEEEauthorrefmark{1}}

\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
Illinois Institute of Technology,
Chicago, Illinois, USA 60616\\
\{xyang56, zzhou\}@hawk.iit.edu, lan@iit.edu}

\IEEEauthorblockA{\IEEEauthorrefmark{2}Mathematics and Computer Science Division, Argonne National Laboratory,
Argonne, IL, USA 60439\\
\{jenkins,rross\}@mcs.anl.gov, mmubarak@anl.gov}
}

% make the title area
\maketitle

\begin{abstract}

Network contention between concurrently running jobs on HPC systems 
is a primary cause of performance variability. 
Optimizing job allocation and avoiding network sharing are 
hence crucial to alleviate the potential performance degradation. 
In order to do these effectively, 
an understanding of the interference among the concurrently running jobs, 
their communication patterns, and the network contention are required. 
In this work, we choose three representative HPC applications from 
DOE Design Forward Project and conduct detailed simulations of 
on a Torus network model to analyze both intra- and inter-job interference. 
By scrutinizing the communication behaviors of these applications, 
we identify relationships between these behaviors and 
the possible interference introduced by different job placement policies. 
Our comprehensive analyses in this work illuminate a path 
towards job-communication-pattern-aware HPC resource allocation.

\end{abstract}

\IEEEpeerreviewmaketitle

\input{intro}
\input{app_study}
\input{research_tool}
%\input{config_study}
\input{interference_study}
\input{discussion}
\input{related_work}
\input{conclusion}

\section*{Acknowledgment}
\label{sec: ack}
The work at Illinois Institute of Technology is supported in part by U.S. National Science Foundation grants CNS-1320125 and CCF-1422009. This work is also supported by the U.S. Department of Energy, Office of Science, Advanced Scientific Computing Research, under Contract DE-AC02-06CH11357.

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv, reference}

\vspace{5\baselineskip}

\begin{framed}
The submitted manuscript has been created by UChicago Argonne, LLC, Operator of Argonne National Laboratory ("Argonne").  Argonne, a U.S. Department of Energy Office of Science laboratory, is operated under Contract No. DE-AC02-06CH11357.  The U.S. Government retains for itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide license in said article to reproduce, prepare derivative works, distribute copies to the public, and perform publicly and display publicly, by or on behalf of the Government.
\end{framed}

\end{document}